---
title: "Course Project 1"
date: "June 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducible Research: Peer Assessment 1

```{r}
library(ggplot2)
library(scales)
library(Hmisc)
library(chron)
```

## Loading and preprocessing the data

1. Load the data (i.e. read.csv())

```{r pressure, echo=FALSE}
data <- read.csv("C:\\Users\\yihe\\Desktop\\Coursera\\Reproducible research\\assignment1\\activity.csv")
```

## Process the data:

```{r}
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
```


## What is mean total number of steps taken per day?


1. Make a histogram of the total number of steps taken each day

```{r}
library(ggplot2)
qplot(total.steps, binwidth=500, xlab="total number of steps taken each day")
```

2. Calculate and report the mean and median total number of steps taken per day

```{r}
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```


## What is the average daily activity pattern?

1. Make a time series plot

```{r}
library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
        geom_line() +
        xlab("5-minute interval") +
        ylab("average number of steps taken")

averages[which.max(averages$steps),] 

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
averages[which.max(averages$steps),]
```


## Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r}
missing <- is.na(data$steps)
# How many missing
table(missing)
```


2. Devise a strategy for filling in all of the missing values in the dataset.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
fill.value <- function(steps, interval) {
        filled <- NA
        if (!is.na(steps))
                filled <- c(steps)
        else
                filled <- (averages[averages$interval==interval, "steps"])
        return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)

total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)


```

4. Make a histogram of the total number of steps taken each day

```{r}
qplot(total.steps, binwidth=500, xlab="total number of steps taken each day")
```

... and Calculate and report the mean and median total number of steps taken per day.

```{r}
mean(total.steps)
median(total.steps)

```



## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two level

```{r}
library(chron)
filled.data$day <- ifelse(is.weekend(filled.data$date),"weekend","weekday")
```

2. Make a panel plot containing a time series plot

```{r}
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
        xlab("5-minute interval") + ylab("Number of steps")
```
